---
title: "Drug"
author: "Zhe Zhao(433707)"
date: "2022/5/26"
output:
  word_document: default
  html_document: default
---
```{r}
library(class)
library(dplyr)
library(readr)
library(ggplot2)
library(caret)
library(verification)
library(corrplot)
library(gmodels)
library(kernlab)
library(tidyverse)
```

## Introduction
The main target of this project is using classification method of machine learning, to build a model explaining whether a particular person consumed cocaine in the last month based on the training sample and generate predictions for all observations from the test sample.

## Import data
Import data and check detailed information.
```{r}
## Read data
drug <- read.csv("C:/Users/zhaoz/Desktop/jiqixuexi/pro/drugs_train.csv")

## get a quick peek at the data
head(drug)

## get dataset detailed infomation
str(drug)

```


## Data Visualization
Plot a boxplot of variable personality_neuroticism
```{r}
boxplot(drug$personality_neuroticism)
```

Plot a boxplot of variable personality_extraversion
```{r}
boxplot(drug$personality_extraversion)
```

Plot a boxplot of variable personality_openness
```{r}
boxplot(drug$personality_openness)
```

Plot a boxplot of variable personality_agreeableness
```{r}
boxplot(drug$personality_agreeableness)
```

Plot a boxplot of variable personality_conscientiousness
```{r}
boxplot(drug$personality_conscientiousness)
```

Plot a boxplot of variable personality_impulsiveness
```{r}
boxplot(drug$personality_impulsiveness)
```

Plot a boxplot of variable personality_sensation
```{r}
boxplot(drug$personality_sensation)
```



Divide the data into training and testing sample.
```{r}
set.seed(123456)
# divide the data into training and testing sample
drug_which_training <- createDataPartition(drug$consumption_cocaine_last_month,
                                            p = 0.7, 
                                            list = FALSE) 

drug_train <-drug[c(drug_which_training),]
drug_test <- drug[-c(drug_which_training),]
## save data
save(list = c("drug_train",
              "drug_test"),
     file = "C:/Users/zhaoz/Desktop/jiqixuexi/pro/drug_train_test.RData")
```
Check if there are any NA values.
```{r}
any(is.na(drug_train))
```

## Data pre-processing
```{r}
# convert all character variables to factors (apart from the 1st column - id)

# create a vector of names
drug_categorical_vars <- 
  sapply(drug_train[, -1], is.character) %>% 
  which() %>% 
  names()

drug_categorical_vars
```

Apply a conversion in a loop and check all factor variables
```{r}
# apply a conversion in a loop
for (variable in drug_categorical_vars) {
  drug_train[[variable]] <- as.factor(drug_train[[variable]])
}
## check all factor variables
glimpse(drug_train)

```

Verify transformation in data frame details, lets see if there are wrong type data in the training sample.
```{r}
# verify transformation in data frame details
str(drug_train)
```

White a function to generate contingency table.
```{r}
# function: generate contingency table
get.contingency.table <- function(dep.var, indep.var, stat.tests=F){
  if(stat.tests == F){
    CrossTable(dep.var, indep.var, digits=1,
               prop.r=F, prop.t=F, prop.chisq=F)
  }else{
    CrossTable(dep.var, indep.var, digits=1,
               prop.r=F, prop.t=F, prop.chisq=F,
               chisq=T, fisher=T)
  }
}
```

## Significant test
For numeric variables, I prefer to use t-test and for others, I use chi-square test and fisher test.

Draw a plot of variable age.
```{r}
plot(drug_train$age)
```
Check levels and label them for analysis.
```{r}
levels(drug_train$age)
drug_train$age<-factor(drug_train$age,
                                           level=c("18-24",
                                                   "25-34",
                                                   "35-44",
                                                   "45-54",
                                                   "55-64",
                                                   "65+"),
                                            labels = c(1:6))
```

Make contingency table and do Chi-square test and fisher test of variable age.
```{r}
get.contingency.table(drug_train$consumption_cocaine_last_month, drug_train$age)
fisher.test(drug_train$consumption_cocaine_last_month, drug_train$age, workspace = 2e7)
chisq.test(drug_train$consumption_cocaine_last_month, drug_train$age)
```

Draw a plot of variable education.
```{r}
plot(drug_train$education)
```
Check levels and label them for analysis.
```{r}
levels(drug_train$education)
drug_train$education<-factor(drug_train$education,
                                           level=c("Left school before 16 years", 
                                                   "Left school at 16 years",
                                                   "Left school at 17 years",                         
                                                   "Left school at 18 years",    
                                                   "Some college or university, no certificate or degree",
                                                   "Professional certificate/ diploma",  
                                                   "University degree",
                                                   "Masters degree",                                      
                                                   "Doctorate degree"),
                                            labels = c(1:9))

```
Make contingency table and do Chi-square test and fisher test of variable education.
```{r}
get.contingency.table(drug_train$consumption_cocaine_last_month, drug_train$education)
fisher.test(drug_train$consumption_cocaine_last_month, drug_train$education, workspace = 2e7)
chisq.test(drug_train$consumption_cocaine_last_month, drug_train$education)

```

Draw a plot of variable country.
```{r}
plot(drug_train$country)
```
Check levels and label them for analysis.
```{r}
drug_train$country<-factor(drug_train$country,
                                           level=c("Australia",
                                                   "Canada", 
                                                   "Ireland",
                                                   "New Zealand",
                                                   "Other",
                                                   "UK",
                                                   "USA"),
                                                labels = c(1:7))
```
Make contingency table and do Chi-square test and fisher test of variable country.
```{r}
get.contingency.table(drug_train$consumption_cocaine_last_month, drug_train$country)
fisher.test(drug_train$consumption_cocaine_last_month, drug_train$country)
chisq.test(drug_train$consumption_cocaine_last_month, drug_train$country)

```

Draw a plot of variable consumption_alcohol.
```{r}
plot(drug_train$consumption_alcohol)
```
Label them for analysis.
```{r}

drug_train$consumption_alcohol<-factor(drug_train$consumption_alcohol,
                                           level=c("never used",
                                                   "used in last day", 
                                                   "used in last week",
                                                   "used in last month",
                                                   "used in last year",
                                                   "used in last decade",
                                                   "used over a decade ago"),
                                                labels = c(1:7))
```
Make contingency table and do Chi-square test and fisher test of variable consumption_alcohol.
```{r}
get.contingency.table(drug_train$consumption_cocaine_last_month, drug_train$consumption_alcohol)
fisher.test(drug_train$consumption_cocaine_last_month, drug_train$consumption_alcohol, workspace = 2e7)
chisq.test(drug_train$consumption_cocaine_last_month, drug_train$consumption_alcohol)
```

Draw a plot of variable ethnicity.
```{r}
plot(drug_train$ethnicity)
```
Label them for analysis.
```{r}

drug_train$ethnicity <- factor(drug_train$ethnicity,
                                           level=c("Asian",
                                                   "Black",
                                                   "Mixed-Black/Asian",
                                                   "Mixed-White/Asian",
                                                   "Mixed-White/Black",
                                                   "Other",
                                                   "White"),
                                                labels = c(1:7))
```
Make contingency table and do Chi-square test and fisher test of variable ethnicity.
```{r}
get.contingency.table(drug_train$consumption_cocaine_last_month, drug_train$ethnicity)
fisher.test(drug_train$consumption_cocaine_last_month, drug_train$ethnicity, workspace = 2e7)
chisq.test(drug_train$consumption_cocaine_last_month, drug_train$ethnicity)
```

Draw a plot of variable consumption_amphetamines.
```{r}
plot(drug_train$consumption_amphetamines)
```
Label them for analysis.
```{r}
drug_train$consumption_amphetamines <- factor(drug_train$consumption_amphetamines,
                                           level=c("never used",
                                                   "used in last day", 
                                                   "used in last week",
                                                   "used in last month",
                                                   "used in last year",
                                                   "used in last decade",
                                                   "used over a decade ago"),
                                                labels = c(1:7))

```
Make contingency table and do Chi-square test and fisher test of variable consumption_amphetamines.
```{r}
get.contingency.table(drug_train$consumption_cocaine_last_month, drug_train$consumption_amphetamines)
fisher.test(drug_train$consumption_cocaine_last_month, drug_train$consumption_amphetamines, workspace = 2e8)
chisq.test(drug_train$consumption_cocaine_last_month, drug_train$consumption_amphetamines)
```

Draw a plot of variable consumption_caffeine.
```{r}
plot(drug_train$consumption_caffeine)
```
Label them for analysis.
```{r}
drug_train$consumption_caffeine <- factor(drug_train$consumption_caffeine,
                                           level=c("never used",
                                                   "used in last day", 
                                                   "used in last week",
                                                   "used in last month",
                                                   "used in last year",
                                                   "used in last decade",
                                                   "used over a decade ago"),
                                                labels = c(1:7))
```
Make contingency table and do Chi-square test and fisher test of variable consumption_caffeine.
```{r}
get.contingency.table(drug_train$consumption_cocaine_last_month, drug_train$consumption_caffeine)
fisher.test(drug_train$consumption_cocaine_last_month, drug_train$consumption_caffeine, workspace = 2e8)
chisq.test(drug_train$consumption_cocaine_last_month, drug_train$consumption_caffeine)
	
```

Draw a plot of variable consumption_chocolate.
```{r}
plot(drug_train$consumption_chocolate)
```
Label them for analysis.
```{r}
drug_train$consumption_chocolate <- factor(drug_train$consumption_chocolate,
                                           level=c("never used",
                                                   "used in last day", 
                                                   "used in last week",
                                                   "used in last month",
                                                   "used in last year",
                                                   "used in last decade",
                                                   "used over a decade ago"),
                                                labels = c(1:7))
```
Make contingency table and do Chi-square test and fisher test of variable consumption_chocolate.
```{r}
get.contingency.table(drug_train$consumption_cocaine_last_month, drug_train$consumption_chocolate)
fisher.test(drug_train$consumption_cocaine_last_month, drug_train$consumption_chocolate, workspace = 2e8)
chisq.test(drug_train$consumption_cocaine_last_month, drug_train$consumption_chocolate)
	
```

Draw a plot of variable consumption_mushrooms.
```{r}
plot(drug_train$consumption_mushrooms)
```
Label them for analysis.
```{r}

drug_train$consumption_mushrooms <- factor(drug_train$consumption_mushrooms,
                                           level=c("never used",
                                                   "used in last day", 
                                                   "used in last week",
                                                   "used in last month",
                                                   "used in last year",
                                                   "used in last decade",
                                                   "used over a decade ago"),
                                                labels = c(1:7))

	
```
Make contingency table and do Chi-square test and fisher test of variable consumption_mushrooms.
```{r}
get.contingency.table(drug_train$consumption_cocaine_last_month, drug_train$consumption_mushrooms)
fisher.test(drug_train$consumption_cocaine_last_month, drug_train$consumption_mushrooms, workspace = 2e8)
chisq.test(drug_train$consumption_cocaine_last_month, drug_train$consumption_mushrooms)
	
```

Draw a plot of variable consumption_nicotine.
```{r}
plot(drug_train$consumption_nicotine)
```
Label them for analysis.
```{r}

drug_train$consumption_nicotine <- factor(drug_train$consumption_nicotine,
                                           level=c("never used",
                                                   "used in last day", 
                                                   "used in last week",
                                                   "used in last month",
                                                   "used in last year",
                                                   "used in last decade",
                                                   "used over a decade ago"),
                                                labels = c(1:7))


```
Make contingency table and do Chi-square test and fisher test of variable consumption_nicotine.
```{r}
get.contingency.table(drug_train$consumption_cocaine_last_month, drug_train$consumption_nicotine)
fisher.test(drug_train$consumption_cocaine_last_month, drug_train$consumption_nicotine, workspace = 2e8)
chisq.test(drug_train$consumption_cocaine_last_month, drug_train$consumption_nicotine)
	
```

Draw a plot of variable consumption_cannabis
```{r}
plot(drug_train$consumption_cannabis)
```
Label them for analysis.
```{r}

drug_train$consumption_cannabis <- factor(drug_train$consumption_cannabis,
                                           level=c("never used",
                                                   "used in last day", 
                                                   "used in last week",
                                                   "used in last month",
                                                   "used in last year",
                                                   "used in last decade",
                                                   "used over a decade ago"),
                                                labels = c(1:7))

```
Make contingency table and do Chi-square test and fisher test of variable consumption_cannabis.
```{r}
get.contingency.table(drug_train$consumption_cocaine_last_month, drug_train$consumption_cannabis)
fisher.test(drug_train$consumption_cocaine_last_month, drug_train$consumption_cannabis, workspace = 2e8)
chisq.test(drug_train$consumption_cocaine_last_month, drug_train$consumption_cannabis)
```

Draw a plot of variable gender.
```{r}
plot(drug_train$gender)
```
Label them for analysis.
```{r}
drug_train$gender <- factor(drug_train$gender,
                                           level=c("female",
                                                   "male"),
                                                labels = c(1:2))
```
Make contingency table and do Chi-square test and fisher test of variable gender.
```{r}
get.contingency.table(drug_train$consumption_cocaine_last_month, drug_train$gender)
fisher.test(drug_train$consumption_cocaine_last_month, drug_train$gender, workspace = 2e8)
chisq.test(drug_train$consumption_cocaine_last_month, drug_train$gender)
```


Draw a plot of variable consumption_cocaine_last_month
```{r}
plot(drug_train$consumption_cocaine_last_month)
```
Label them for analysis.
```{r}
drug_train$consumption_cocaine_last_month <- factor(drug_train$consumption_cocaine_last_month,
                                           level=c("Yes",
                                                   "No"),
                                                labels = c(0:1))
```

For numeric variables, i used t-test to check the significance.

Do t-test of variable personality_neuroticism.
```{r}
t.test(drug_train$personality_neuroticism ~ drug_train$consumption_cocaine_last_month)
```
Do t-test of variable personality_extraversion.
```{r}
t.test(drug_train$personality_extraversion  ~ drug_train$consumption_cocaine_last_month)
```
Do t-test of variable personality_openness.
```{r}
t.test(drug_train$personality_openness ~ drug_train$consumption_cocaine_last_month)
```
Do t-test of variable personality_agreeableness.
```{r}
t.test(drug_train$personality_agreeableness ~ drug_train$consumption_cocaine_last_month)
```
Do t-test of variable personality_conscientiousness.
```{r}
t.test(drug_train$personality_conscientiousness ~ drug_train$consumption_cocaine_last_month)
```
Do t-test of variable personality_impulsiveness.
```{r}
t.test(drug_train$personality_impulsiveness ~ drug_train$consumption_cocaine_last_month)
```
Do t-test of variable personality_sensation.
```{r}
t.test(drug_train$personality_sensation ~ drug_train$consumption_cocaine_last_month)
```

After these test, we chose below variables: "age", "gender", "country", "ethnicity", "consumption_alcohol","consumption_amphetamines","consumption_cannabis","consumption_mushrooms","consumption_nicotine", "consumption_cocaine_last_month","personality_neuroticism", "personality_agreeableness","personality_conscientiousness", "personality_impulsiveness", "personality_sensation"

Creat a new dataframe with selected variables.
```{r}
load("C:/Users/zhaoz/Desktop/jiqixuexi/pro/drug_train_test.RData")
drug_train_pro <- drug_train[,c("age", "gender", "country", "ethnicity", "consumption_alcohol", 
                              "consumption_amphetamines","consumption_cannabis",
                              "consumption_mushrooms","consumption_nicotine", "consumption_cocaine_last_month",
                              "personality_neuroticism", "personality_agreeableness", 
                              "personality_conscientiousness", "personality_impulsiveness", "personality_sensation")]
```

And for testing sample.
```{r}
drug_test <- drug_test[,c("age", "gender", "country", "ethnicity", "consumption_alcohol", 
                              "consumption_amphetamines","consumption_cannabis",
                              "consumption_mushrooms","consumption_nicotine", "consumption_cocaine_last_month",
                              "personality_neuroticism", "personality_agreeableness", 
                              "personality_conscientiousness", "personality_impulsiveness", "personality_sensation")]
```


## KNN model

Firstly, we use drug_train_pro this new dataset to check the performance of the model.

Define the training control, I use 5-fold cross validation.
```{r}
## set “k”s
different_k <- data.frame(k = seq(1, 99, 4))
## use 5-fold cross validation
ctrl_cv5 <- trainControl(method = "cv",
                         number = 5)
```

## Model 1

```{r}
## run the training
set.seed(2222)

drug_train_pro_knn_tuned <- 
  train(consumption_cocaine_last_month ~ ., 
        data = drug_train_pro,
        method = "knn",
        # validation used!
        trControl = ctrl_cv5,
        # parameters to be compared
        tuneGrid = different_k)

## now validation is applied to EVERY SINGLE value of k.

## lets check the results

drug_train_pro_knn_tuned
plot(drug_train_pro_knn_tuned)
```

We can see that from the perspective of AUC the best value of k = 97. 
Value of Accuracy is good, the best accuracy obtained >> k = 97 and the accuracy seems to increase with k.

## Model 2 
In the case of cross-validation, should apply transformations of input data in the train() function, I use preProcess = "range".
```{r}
set.seed(222)

drug_train_pro_knn_tuned_scaled <- 
  train(consumption_cocaine_last_month ~ ., 
        data = drug_train_pro,
        method = "knn",
        # validation used!
        trControl = ctrl_cv5,
        # parameters to be compared
        tuneGrid = different_k,
        # data transformation
        preProcess = c("range"))


drug_train_pro_knn_tuned_scaled
plot(drug_train_pro_knn_tuned_scaled)
```
We can see that from the perspective of AUC the best value of k = 9.
Value of Accuracy is very similar to above.


Check which value of k gives the highest AUC.
```{r}
ctrl_cv5a <- trainControl(method = "cv",
                          number = 5,
                          # probabilities of each level predicted in cross-validation
                          classProbs = TRUE,
                          # summary function that includes ROC
                          summaryFunction = twoClassSummary)

```

## Model 3
In the case of cross-validation, should apply transformations of input data in the train() function, I use preProcess = "range". And I also use validation with probabilities and twoClassSummary
```{r}
set.seed(2222)
drug_train_pro_knn_tuned_scaled2 <- 
  train(consumption_cocaine_last_month ~ ., 
        data = drug_train_pro,
        # knn
        method = "knn",
        # validation used - with probabilities and twoClassSummary
        trControl = ctrl_cv5a,
        # parameters to be compared
        tuneGrid = different_k,
        preProcess = c("range"),
        metric = "ROC")

drug_train_pro_knn_tuned_scaled2
plot(drug_train_pro_knn_tuned_scaled2)

```



Now run the model for testing sample and check.
Testing sample is names “drug_test”, separated from “drug_train”.
```{r}
drug_test$predicted <- drug_train_pro_knn_tuned_scaled2 %>% predict(drug_test)
```

Lets built a confusion matrix table.
```{r}
# confusion matrix table 
confusion_table = table(drug_test[,'predicted'], drug_test[,'consumption_cocaine_last_month'])
confusion_table
```

Definite of TP,TN,FP,FN.
```{r} 
## Definition of TP,TN,FP,FN
confusion_table[1,1] = 'TN'
confusion_table[1,2] = 'FN'
confusion_table[2,1] = 'FP'
confusion_table[2,2] = 'TP'
## output
confusion_table
```

Write a function to calculate accuracy.
```{r}
## function
et_accuracy <- function(df, predicted, actual){
  confusion_table = table(drug_test[,'predicted'], drug_test[,'consumption_cocaine_last_month'])
  TP = confusion_table[2,2]
  TN = confusion_table[1,1]
  FN = confusion_table[1,2]
  FP = confusion_table[2,1]
  accuracy = round((TP + TN) / sum(TP,FP,TN,FN), 2)
  return(accuracy)
}
```

Use this function to check the accuracy of testing sample!
```{r}
## accuracy value 
score = data.frame(accuracy=et_accuracy(drug_test,'predicted','consumption_cocaine_last_month'))

## output
score
                   
```
Seems good!

## Conclusion
I also used SVM and decision tree, the best method is KNN, because the Accuracy value is highest.

